### 2023-10-05 proposal/proposal.tex
added sections/outline to proposal

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: }
\author{Luke Noel\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}

\paragraph{Specific Aims}

\paragraph{Data}

\paragraph{Research Design and Methods}

\paragraph{Discussion}

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}