### 2023-10-10 proposal/proposal.tex
finished intro

\documentclass[12pt]{article}

%% preamble:
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: Evaluating the Impact of 2023 MLB Rule Changes on Performance
 Metrics and Strategies in Major League Baseball}
\author{Luke Noel\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
Major League Baseball (MLB) is a dynamic sport that continually evolves through 
rule changes aimed at enhancing game pace, player safety, and fan engagement. In
the 2023 MLB season, significant rule changes were introduced, including the 
implementation of a pitch clock, outlawing the defensive shift, and larger 
bases. This research proposal outlines a comprehensive study that aims to
analyze the impact of these rule changes on various performance metrics in
baseball, including but not limited to stolen bases, pace of play,
and batting averages.

\paragraph{Specific Aims}

\paragraph{Data}

\paragraph{Research Design and Methods}

\paragraph{Discussion}

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}