### 2023-10-10 proposal/proposal.tex
finished aims, data, and methods sections of proposal

\documentclass[12pt]{article}

%% preamble:
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: Evaluating the Impact of 2023 MLB Rule Changes on Performance
 Metrics and Strategies in Major League Baseball}
\author{Luke Noel\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
Major League Baseball (MLB) is a dynamic sport that continually evolves through 
rule changes aimed at enhancing game pace, player safety, and fan engagement. In
the 2023 MLB season, significant rule changes were introduced, including the 
implementation of a pitch clock, the ban of the defensive shift, and larger 
bases. This research proposal outlines a comprehensive study that aims to
analyze the impact of these rule changes on various performance metrics in
baseball, including but not limited to stolen bases, pace of play,
and batting averages.

\paragraph{Specific Aims}
The specific aims of this paper are to investigate the impact of the 2023 MLB
rule changes on various performance metrics and to assess whether these changes
had significant effects. Some research questions I will address are:

\begin{enumerate}
  \item How did the 2023 MLB rule changes influence the frequency and success 
  rate of stolen base attempts?
  \item Were there statistically significant changes in time of play before and
   after the rule changes?
  \item Did the rule changes have a noticeable impact on batting averages and 
  other offensive metrics?
  \item What are the effect sizes associated with the observed differences
   in these metrics?
\end{enumerate}

\paragraph{Data}
The data for this research will consist of comprehensive statistics for the 2023
MLB season, including game logs, and player/team/league performance metrics.
This data will then be compared and contrasted with the same data but 
from previous seasons where these rule changes were not implemented yet. There
are various sources where the data will be obtained from: \href{https://www.espn.com/mlb/stats/team/_/season/2023/seasontype/2}{ESPN}
and \href{https://www.baseball-reference.com/}{Baseball Reference} will give
league, team, and player averages across seasons along with game duration.
\href{https://baseballsavant.mlb.com/league?season=2022#statcastHitting}{Baseball Savant}
will be useful for providing advanced statistics.

\paragraph{Research Design and Methods}
To address the research questions, a combination of statistical methods will be
used. This may include two sample hypothesis tests, regression analysis, and
various data visualization techniques to truly show the impact of the rule
changes. Each performance metric such as batting average, home runs, and stolen
bases will be analyzed separately to determine whether the rule changes had a
statistically significant impact.

\paragraph{Discussion}

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}